---
title: "Coursera´s PML Assignment writeup"
author: "Farrukh Shahzad"
date: "26th April 2015"
---

##Summary

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement? a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).  

###Data 

The training data for this project are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here: 

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv


## Data Pre-processing  

**Performed steps**:  
* Get the system ready loading dependencies and the data  
* Remove strings `#DIV/0!` and `NA` values    

```{r Dependencies and data, cache=TRUE}
library(Hmisc)
library(caret)
library(corrplot)
library(randomForest)
library(foreach)
library(doParallel)
set.seed(1980)
options(warn=-1)

training_data <- read.csv("pml-training.csv", header = TRUE, na.strings = c("NA", "#DIV/0!", ""))
evaluation_data <- read.csv("pml-testing.csv", header = TRUE, na.strings = c("NA", "#DIV/0!", ""))

corMat <- cor(training_data[, -54])
corrplot(corMat, order = "FPC", method = "color", type = "lower", tl.cex = 0.8, 
         tl.col = rgb(0, 0, 0))
```

Set columns from the eighth to the end to be numeric.  

```{r Arranging columns}
for(i in c(8:ncol(training_data)-1)) {training_data[,i] = as.numeric(as.character(training_data[,i]))}

for(i in c(8:ncol(evaluation_data)-1)) {evaluation_data[,i] = as.numeric(as.character(evaluation_data[,i]))}
```

There are some almost-totally blank columns in the data that do not contribute well to the prediction, so the feature dataset is made to only include complete columns.  Also removing user name, timestamps and windows.  


##Machine Learning

Determine and display out feature set.

```{r Feature Set}
feature_set <- colnames(training_data[colSums(is.na(training_data)) == 0])[-(1:7)]
model_data <- training_data[feature_set]
feature_set
```

The model data is now built built from the feature set.

```{r Model Dataset}
idx <- createDataPartition(y=model_data$classe, p=0.75, list=FALSE )
training <- model_data[idx,]
testing <- model_data[-idx,]
```

Build 5 random forests with 150 trees each using parallel processing to build this model.

```{r Building the forests, cache=TRUE,echo=TRUE}
registerDoParallel()
x <- training[-ncol(training)]
y <- training$classe

rf <- foreach(ntree=rep(150, 6), .combine=randomForest::combine, .packages='randomForest') %dopar% {
randomForest(x, y, ntree=ntree) 
}
```

##Predicted Results

Provide error reports for both training and test data.  
```{r Error reports, cache=TRUE,echo=TRUE}
predictions1 <- predict(rf, newdata=training)
confusionMatrix(predictions1,training$classe)


predictions2 <- predict(rf, newdata=testing)
confusionMatrix(predictions2,testing$classe)
```

##Conclusions and Test Data Submit


The confusion matrix shows a fairly accurate model. The test data was around 99% 
accurate so it was expected that most of the submitted test cases were correct. In the end all of them showed to be correct.

###Use the code for file creation for all 20 answers

```{r File submission}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}


x <- evaluation_data
x <- x[feature_set[feature_set!='classe']]
answers <- predict(rf, newdata=x)
answers

##  [1] B A C A A E D B A A B C B A E E A B B B
## Levels: A B C D E

pml_write_files(answers)
```
